Time Is Running Out

### 2023-05-09 abstract.tex
ch3 update

In a society where social media is so ubiquitous present, the privacy concerns
around them are more relevant than ever. While developing applications, privacy
laws and concerns must be taken into account. But this does not mean all these
platforms that where built whit those in mind are bulletproof. In a lot of
applications is it still possible to find vulnerabilities in the system, with
the possibility of rather unpleasant consequences. During this thesis, the main
focus will be on the privacy policies of fitness trackers. Fitness trackers are
platforms which store and display data related to sport activities. These can
be shared with other users, to show your achievements, and possibly motivating
others to exercise as well. This data may include heart rate, GPS-locations,
etc. Some pieces could potentially be more privacy-sensitive than others. The
relevant data to study in this thesis are GPS-locations and GPS-related data
(like speed, distance, \ldots). A great concern about sharing GPS-data, is
potentially sharing locations you would rather keep private, for example your
home location. Sharing full GPS data of your activities could leak this
location.

Most fitness tracking networks are aware of this danger and implement a series
of countermeasures to prevent leaking this sensitive information.
Countermeasures are coming however with a cost, namely a (slightly) worse user
experience. From the perspective of the developers of the fitness trackers, a
trade-off is consistently being made between privacy and user experience. On
most platforms like Strava, Garmin, are similar basic privacy features
implemented. These are features like hiding activities, or only sharing
activities with your followers. Another commonly used countermeasure is a
mechanism known as an \textit{EPZ} (Endpoint Privacy Zone).

An \textit{EPZ} is a circle or polygon drawn around a certain sensitive
location. The circular EPZ's will de drawn using a radius chosen by the user,
and a center which is a random point in the area of around the sensitive
location. This center can't be further than 70\% of the radius away from this
sensitive location. When this zone is generated, the end and beginning of the
trajectory followed which pass through this zone will be hidden for other
users.

Most EPZ implementation are not perfect in assuring privacy. While hiding these
parts, other useful information is not being hidden or adapted to this sort of
cloaking. During this thesis, the goal is to retrieve sensitive locations. This
can be achieved by using the total times and distances of the activities.
Previous research showed that it is possible to retrieve sensitive locations
using the total distance combined with the street map of the area. These
attacks are called \textit{inference attacks}. The distance traveled inside the
EPZ can be inferred using the total distance given by the API, and the distance
traveled outside of the EPZ (this is the visible distance on the map). Using
the distance traveled inside of the EPZ, a route can be constructed and mapped
onto the street plan. If all the possible routes are considered, multiple
possible locations are found. If this is repeated for different activities,
with different points where the EPZ is being entered, only one point will
remain (in the best case). This would then be the sensitive location.

This thesis investigates the possibilities of such inference attacks using data
other than the distance as a base. In our implementation, the speed and tempo
of the activities will mainly be the used, in combination with the
GPS-locations. This method will consist of three parts. First, the average
speed and the total duration will be used to calculate the total distance.
Second, the GPS-points will be used to calculate the distance traveled outside
of the \ac{EPZ}. In order to do this effectively, smoothing and map snapping
strategies need to be tested out to get the best possible results. These two
values can be used in the third step to execute the interference attack. The
results of this attack will be compared with the results of the previous
implementations of this sort of attack.

This attack is successful in some cases. With the correct tuning of the
parameters of the smoothing algorithm, a succes rate of 75\% can be achieved.
This is lower than previous implementations of this attack, which was as
expected considering the type of data that is being used. The GPS locations in
particular are not always accurate. And because there are so many GPS-points
are needed for these calculations, small deviations on every point result in a
large deviation on the calculated distance. But the main conclusion is that
this attack is possible, with a reasonable succes rate.

>>>>

In a society where social media is so ubiquitous present, the privacy concerns
around them are more relevant than ever. While developing applications, privacy
laws and concerns must be taken into account. But this does not mean all these
platforms that where built whit those in mind are bulletproof. In a lot of
applications it is still possible to find vulnerabilities in the system, with
the possibility of rather unpleasant consequences. During this thesis, the main
focus will be on the privacy policies of fitness trackers. Fitness trackers are
platforms which store and display data related to sport activities. These can
be shared with other users, to show your achievements, and possibly motivating
others to exercise as well. This data may include heart rate, GPS-locations,
etc. Some pieces could potentially be more privacy-sensitive than others. The
relevant data to study in this thesis are GPS-locations and GPS-related data
(like speed, distance, \ldots). A great concern about sharing GPS-data, is
potentially sharing locations you would rather keep private, for example your
home location. Sharing full GPS data of your activities could leak this
location.

Most fitness tracking networks are aware of this danger and implement a series
of countermeasures to prevent leaking this sensitive information.
Countermeasures are coming however with a cost, namely a (slightly) worse user
experience. From the perspective of the developers of the fitness trackers, a
trade-off is consistently being made between privacy and user experience. On
most platforms like Strava, Garmin, similar basic privacy features are
implemented. These are features like hiding activities, or only sharing
activities with your followers. Another commonly used countermeasure is a
mechanism known as an \textit{EPZ} (Endpoint Privacy Zone).

An \textit{EPZ} is a circle or polygon drawn around a certain sensitive
location. The circular EPZ's will be drawn using a radius chosen by the user,
and a center which is a random point in the area of around the sensitive
location. This center can't be further than 70\% of the radius away from this
sensitive location. When this zone is generated, the end and beginning of the
trajectory followed which pass through this zone will be hidden for other
users.

Most EPZ implementation are not perfect in assuring privacy. While hiding these
parts, other useful information is not being hidden or adapted to this sort of
cloaking. During this thesis, the goal is to retrieve sensitive locations. This
can be achieved by using the total times and distances of the activities.
Previous research showed that it is possible to retrieve sensitive locations
using the total distance combined with the street map of the area. These
attacks are called \textit{inference attacks}. The distance travelled inside
the EPZ can be inferred using the total distance given by the API, and the
distance travelled outside of the EPZ (this is the visible distance on the
map). Using the distance travelled inside of the EPZ, a route can be
constructed and mapped onto the street plan. If all the possible routes are
considered, multiple possible locations are found. If this is repeated for
different activities, with different points where the EPZ is being entered,
only one point will remain (in the best case). This would then be the sensitive
location.

This thesis investigates the possibilities of such inference attacks using data
other than the distance as a base. In our implementation, the speed and tempo
of the activities will be used, in combination with the GPS-locations. This
method will consist of three parts. First, the average speed and the total
duration will be used to calculate the total distance. Second, the GPS-points
will be used to calculate the distance travelled outside of the \ac{EPZ}. In
order to do this effectively, smoothing and map snapping strategies need to be
tested out to get the best possible results. These two values can be used in
the third step to execute the interference attack. The results of this attack
will be compared with the results of the previous implementations of this sort
of attack.

This attack is successful in some cases. With the correct tuning of the
parameters of the smoothing algorithm, a succes rate of 75\% can be achieved.
This is lower than previous implementations of this attack, which was as
expected considering the type of data that is being used. The GPS locations in
particular are not always accurate. And because there are so many GPS-points
needed for these calculations, small deviations on every point result in a
large deviation on the calculated distance. But the main conclusion is that
this attack is possible, with a reasonable succes rate.

### 2023-05-09 hfdst1.tex
ch3 update

\section{Situering}
Sociale media is zo goed als niet meer weg te denken uit het huidige moderne
leven. Over de jaren heen zijn er verschillende definities gegeven. in In het
werk van Howard en Park wordt sociale media gedefinieerd als de infrastructuur
en tools om content te maken en te verspreiden\cite{PhilipsAndParks}. Deze
definitie is erg ruim, en vertakt zich dus in heel wat facetten, waaronder
sociale netwerken, media sharing networks, etc. Maar ook de fitnesstrackers.
Deze opkomst van nieuwe media brengt echter ook onbedoelde maar significante
privacy bezorgdheden met zich mee.

De focus in deze dissertatie ligt op privacy binnen fitnesstrackers, meer
specifiek platformen die \ac{gps}-locaties gebruiken, zoals Strava, Nike Run Club,
etc. Dit zijn platformen waar personen sportactiviteiten zoals lopen, fietsen,
wandelen,\ldots kunnen delen met elkaar. Het algemene concept is hierbij dat
wanneer je een sportactiviteit uitvoert, je deze voor je volgers en vrienden
beschikbaar maakt. De sportactiviteiten zullen natuurlijk bepaalde gegevens
bevatten die zichtbaar zijn voor die andere gebruikers,
Figuur~\ref{fig:activityExample} geeft bijvoorbeeld weer hoe Strava de afstand,
bewegingstijd, en natuurlijk de \ac{gps}-locaties deelt. Vele van deze gegevens
hebben direct of indirect een negatieve impact op de privacy van de user. Deze
negatieve gevolgen komen dan vooral in de vorm van het onbedoeld vrijgeven van
\textit{gevoelige locaties}. Onder het concept van een gevoelige locatie vallen
heel wat beschrijvingen. Een algemene beschrijving kan zijn, een locatie die
geografische informatie deelt die negatieve gevolgen kan hebben, en die je dus
liever niet deel. In het kader van dit onderzoek, zal dit dan gaan over start
en eindlocaties van activiteiten. Dit kan gaan over woonplaatsen, wat kan
leiden tot o.a.\ stalking. Alsook locaties waar sportmateriaal wordt
opgeborgen. Er zijn gevallen bekend van fietsdieven die Strava gebruiken om
fietsen te kunnen lokaliseren\cite{Sportapp72:online}\cite{Cyclistw89:online}.
Grootschaligere voorbeelden die zeker het vermelden waard zijn, zijn de
gevallen waarbij geheime militaire basissen ontdekt worden door het bestuderen
van de heatmap\cite{Fitnesst33:online}.
\begin{figure}
    \centering
    \includegraphics[width=0.5\linewidth]{fig/VoorbeeldActiviteiten/VoorbeeldActiviteit_Cropped.png}
    \caption{Voorbeeldactiviteit Strava}\label{fig:activityExample}
\end{figure}

Deze platformen implementeren elk manieren om de privacy van de users te
verbeteren. De meest eenvoudige te bedenken is misschien wel de mogelijkheid om
activiteiten te verbergen voor een selectie van personen (bv.\ iedereen die
geen volger is). Zo kunnen enkel de mensen die de gebruiker expliciet toelaat
activiteiten bekijken. Een complexer alternatief is het gebruik van \acp{EPZ}.
Hierbij wordt de weergegeven route voor de persoon die meekijkt gedeeltelijk
verborgen. Er wordt als het ware een deel van de route afgekapt. De echte
begin- en eindpunten zullen binnenin het afgekapte deel liggen. Er zullen
nieuwe punten worden gegenereerd worden, op de rand van de cirkel, die voor de
externe waarnemer het begin en einde zullen voorstellen. Het begin- en
eind-deel van de route wordt dus onzichtbaar voor de andere gebruikers. Door de
aanwezigheid van al deze pogingen tot privacyverbeteringen valt op dat de
ontwikkelaars van de platformen erg bewust zijn van de mogelijke gevaren.
Echter is er een afweging te maken bij de implementatie tussen de bruikbaarheid
van het platform, en de privacy van de eindgebruiker. Hoe meer data wordt
vrijgegeven, hoe groter de kans om mogelijk gevoelige info wordt meegegeven.
Aan de andere kant, bij het weglaten van informatie gaat de
gebruiksvriendelijkheid en de aanwezigheid van nuttige data van het platform
serieus achteruit.

>>>>

\section{Situering}
Sociale media is zo goed als niet meer weg te denken uit het huidige moderne
leven. Over de jaren heen zijn er verschillende definities gegeven. In het werk
van Howard en Park wordt sociale media gedefinieerd als de infrastructuur en
tools om content te maken en te verspreiden\cite{PhilipsAndParks}. Deze
definitie is erg ruim, en vertakt zich dus in heel wat facetten, waaronder
sociale netwerken, media sharing networks, etc. Maar ook de fitnesstrackers.
Deze opkomst van nieuwe media brengt echter ook onbedoelde maar significante
privacy bezorgdheden met zich mee.

De focus in deze dissertatie ligt op privacy binnen fitnesstrackers, meer
specifiek platformen die \ac{gps}-locaties gebruiken, zoals Strava, Nike Run
Club, etc. Dit zijn platformen waar personen sportactiviteiten zoals lopen,
fietsen, wandelen,\ldots kunnen delen met elkaar. Het algemene concept is
hierbij dat wanneer je een sportactiviteit uitvoert, je deze voor je volgers en
vrienden beschikbaar maakt. De sportactiviteiten zullen natuurlijk bepaalde
gegevens bevatten die zichtbaar zijn voor die andere gebruikers,
Figuur~\ref{fig:activityExample} geeft bijvoorbeeld weer hoe Strava de afstand,
bewegingstijd, en natuurlijk de \ac{gps}-locaties deelt. Vele van deze gegevens
hebben direct of indirect een negatieve impact op de privacy van de user. Deze
negatieve gevolgen komen dan vooral in de vorm van het onbedoeld vrijgeven van
\textit{gevoelige locaties}. Onder het concept van een gevoelige locatie vallen
\textit{gevoelige locaties}. Onder het concept van een gevoelige locatie vallen
heel wat beschrijvingen. Een algemene beschrijving kan zijn, een locatie die
geografische informatie deelt die negatieve gevolgen kan hebben, en die je dus
liever niet deelt. In het kader van dit onderzoek zal dit gaan over start en
eindlocaties van activiteiten. Dit kan gaan over woonplaatsen, wat kan leiden
tot o.a.\ stalking. Alsook locaties waar sportmateriaal wordt opgeborgen. Er
zijn gevallen bekend van fietsdieven die Strava gebruiken om fietsen te kunnen
lokaliseren\cite{Sportapp72:online}\cite{Cyclistw89:online}. Grootschaligere
voorbeelden die zeker het vermelden waard zijn, zijn de gevallen waarbij
geheime militaire basissen ontdekt worden door het bestuderen van de
heatmap\cite{Fitnesst33:online}.
\begin{figure}
    \centering
    \includegraphics[width=0.5\linewidth]{fig/VoorbeeldActiviteiten/VoorbeeldActiviteit_Cropped.png}
    \caption{Voorbeeldactiviteit Strava}\label{fig:activityExample}
\end{figure}

Deze platformen implementeren elk manieren om de privacy van de users te
verbeteren. De meest eenvoudige te bedenken is misschien wel de mogelijkheid om
activiteiten te verbergen voor een selectie van personen (bv.\ iedereen die
geen volger is). Zo kunnen enkel de mensen die de gebruiker expliciet toelaat
activiteiten bekijken. Een complexer alternatief is het gebruik van \acp{EPZ}.
Hierbij wordt de weergegeven route voor de persoon die meekijkt gedeeltelijk
verborgen. Er wordt als het ware een deel van de route afgekapt. De echte
begin- en eindpunten zullen binnenin het afgekapte deel liggen. Er zullen
nieuwe punten worden gegenereerd, op de rand van de cirkel, die voor de externe
waarnemer het begin en einde zullen voorstellen. Het begin- en eind-deel van de
route wordt dus onzichtbaar voor de andere gebruikers. Door de aanwezigheid van
al deze pogingen tot privacyverbeteringen valt op dat de ontwikkelaars van de
platformen erg bewust zijn van de mogelijke gevaren. Echter is er een afweging
te maken bij de implementatie tussen de bruikbaarheid van het platform, en de
privacy van de eindgebruiker. Hoe meer data wordt vrijgegeven, hoe groter de
kans op mogelijk gevoelige info wordt meegegeven. Aan de andere kant, bij het
weglaten van informatie gaat de gebruiksvriendelijkheid en de aanwezigheid van
nuttige data van het platform serieus achteruit.
